Preventing Runtime Errors of Redis at Compile Time

### 2015-11-20 paper.tex
fix all \todos

\todo{scm}{Cite some references for Redis?}

Redis is an open source, in-memory data structure store, often used as
database, cache and message broker. A Redis datatype can be think of as a set of key-value pairs, where each value is associated with a binary-safe string key to identify and manipulate with. Redis supports many different kind of values, such as strings, hashes, lists, and sets, etc, and provides a collection of of atomic \emph{commands} to manipulate these values.

@@

Many third party library exist to allow general purpose programmings
to access Redis databases through its TCP protocol. The most popular
such library for Haskell is Hedis.
%
\todo{scm}{Citation for Hedis?}
%
The following program is how the previous example looks like in
Hedis:

@@

The function \mintinline{haskell}{sadd} takes a key and a list of values as arguments, and
 returns an \mintinline{haskell}{Integer}, wrapped in
 \mintinline{haskell}{Either} to indicate possible failures in the context
 \mintinline{haskell}{Redis}\footnotemark of command execution.
\todo{scm}{Explain Reply in one sentence?}

@@

\say{All binary strings are equal, but some binary strings are more equal than others.}

\todo{scm}{Is this a quote? If so please cite the source. If not,
we shouldn't use the double quotes.}

@@

\paragraph{The Cause} These problems arise from the absence of type checking,
 with respects to \textbf{the type of a value that associates with a key}.

\todo{scm}{say a bit more.}

@@

Haskell makes it easy to build and use Domain Specific Languages (DSLs),
 and Hedis can be regarded as one of them. What makes Hedis special is that,
 it has \emph{variable bindings} (between keys and values), but with very
 little or no semantic checking, neither dynamically nor statically.

We began with making Hedis a dynamically checked embedded DSL, and implemented a
 runtime type checker that keeps track of everything. But then we found that
 things can be a lot easier, by leveraging the host language's type checker,
 which also make it statically type-checked!

@@

@@

We have shown that how a embedded DSL with variable-bindings could be expressed
 in types and statically checked. With recent extensions to the language,
 Haskell is becoming more and more capable of dependently-typed programming, and
 it's not that scary as many (including me) would have thought.

>>>>

Redis\footnote{\url{https://redis.io}} is an open source, in-memory data structure store, often used as
database, cache and message broker. A Redis datatype can be think of as a set of key-value pairs, where each value is associated with a binary-safe string key to identify and manipulate with. Redis supports many different kind of values, such as strings, hashes, lists, and sets, etc, and provides a collection of of atomic \emph{commands} to manipulate these values.

@@

Many third party library exist to allow general purpose programmings
to access Redis databases through its TCP protocol. The most popular
such library for Haskell is Hedis\footnote{\url{https://hackage.haskell.org/package/hedis}}.

The following program is how the previous example looks like in
Hedis:

@@

The function \mintinline{haskell}{sadd} takes a key and a list of values as
 arguments, and returns an \mintinline{haskell}{Integer} on success, or
 returns a \mintinline{haskell}{Reply}, a low-level representation of replies
 from the Redis server, in case of failues. All wrapped in
 \mintinline{haskell}{Redis}\footnotemark, the context of command execution.

@@

All binary strings are equal, but some binary strings are more equal than others.

@@

\paragraph{The Cause} Every key is associated with a value, and every value has
 it's own type. But Most commands in Redis only work with a certain type of
 value. When a command is used on a wrong type of key, a runtime error occurs.
 The problems illustrated above arise from the absence of type checking, with
 respects to \textbf{the type of a value that associates with a key}.
 These problems could have been avoided, if we could know the type every key
 associates with in advance, and prevent programs with invalid commands from
 executing.

@@

Haskell makes it easy to build and use Domain Specific Languages (DSLs),
 and Hedis can be regarded as one of them. What makes Hedis peculiar is that,
 it has \emph{variable bindings} (between keys and values), but with very
 little or no semantic checking, neither dynamically nor statically.

We began with making Hedis a dynamically checked embedded DSL, and implemented a
 runtime type checker that keeps track of types of each variable. But then we
 found that things can be a lot easier, by leveraging the host language's type
 checker. We encode variable bindings with \emph{type-level lists} and
 \emph{strings}, and control the effects on the bindings with
 \emph{indexed monad}. In contrast to the former approach, we \textbf{embedded}
 our type checker into Haskell's type system, without having to build a
 \textbf{standalone} one on the term level.

@@

Notice that \textbf{not all Hedis programs can be typechecked} (even if they
 might turn out to be type safe). We opted for type safety instead of
 expressiveness.

@@

We have demonstrated how a DSL with variable-bindings could be embedded in types.
With more and more extensions added to the language, Haskell is gradually
 becoming a dependently-typed language,\footnote{Why not be dependently typed? \url{http://stackoverflow.com/questions/12961651/why-not-be-dependently-typed}}
 but it's not that dreadful as many (including us) would have thought.
